### 2019-09-04 initial/notes.tex
validated loss of stability
stash some notes

Now, a slightly more subtle question: is it possible for CS2 to still exist but
lose stability? Consider that for some $\epsilon\lesssim \epsilon_c$, the
equilibria lie at $\phi_{2, 4} = \pi/2 \pm \Delta \phi$, where $\epsilon_c -
\epsilon = \eta \sin I \frac{\Delta \phi^2}{2}$, then we consider perturbation
about these equilibria $\phi_{2, 4} = \pi/2 \pm \Delta \phi + \delta \phi$. The
linearized equations then read
\begin{subequations}
    \begin{align}
        \dot{\phi} &\approx -\delta \mu \mp \eta^2 \sin I \cos I \delta \phi,\\
        \dot{\mu} &\approx \pm\eta \sin I \Delta \phi \delta \phi.
    \end{align}
\end{subequations}
The eigenvalue equation for this first-order system reads
\begin{equation}
    \lambda^2 \pm \eta^2 \sin I \cos I \pm \eta \sin I \Delta \phi = 0.
\end{equation}
This has real solutions if
\begin{equation}
    \p*{\eta^2 \sin I \cos I}^2 \mp \eta \sin I \Delta \phi > 0.
\end{equation}
If $\epsilon \sim \epsilon_c / 2$, i.e.\ not too close to critical, then
$\Delta \phi^2 \sim 1$, so except for $\epsilon_c - \epsilon \ll \epsilon_c$,
the second term dominates the first and we have one real and one imaginary root,
corresponding to the two CSs. Otherwise, we can identify that for
\begin{align*}
    \Delta \phi &\leq \eta^3 \sin I \cos^2 I,\\
    \epsilon_c - \epsilon &\leq \frac{\eta^7 \sin^3 I \cos^4 I}{2},
\end{align*}
that CS2 becomes unstable.

>>>>

In fact, we can really do better: just let $\sin \phi = \frac{\epsilon\p*{1 -
z_0^2}}{\eta \sin I}$, then $\mu = \frac{\eta \cos I}{1 - \eta \sin I \cos \phi
\frac{1}{\sqrt{1 - z_0}}}$ where $z_0 = \eta \cos I$ is the zeroth order
estimate. This is what is plotted in the aforementioned plot.

Now, a slightly more subtle question: is it possible for CS2 to still exist but
lose stability? Consider that for some $\epsilon\lesssim \epsilon_c$, the
equilibria lie at $\phi_{2, 4} = \pi/2 \pm \Delta \phi$, where $\epsilon_c -
\epsilon = \eta \sin I \frac{\Delta \phi^2}{2}$, then we consider perturbation
about these equilibria $\phi_{2, 4} = \pi/2 \pm \Delta \phi + \delta \phi$. The
linearized equations then read
\begin{subequations}
    \begin{align}
        \dot{\phi} &\approx -\delta \mu \mp \eta^2 \sin I \cos I \delta \phi,\\
        \dot{\mu} &\approx \pm\eta \sin I \Delta \phi \delta \phi.
    \end{align}
\end{subequations}
The eigenvalue equation for this first-order system reads
\begin{equation}
    \lambda^2 \pm \eta^2 \sin I \cos I \pm \eta \sin I \Delta \phi = 0.
\end{equation}
This has real solutions if
\begin{equation}
    \p*{\eta^2 \sin I \cos I}^2 \mp \eta \sin I \Delta \phi > 0.
\end{equation}
If $\epsilon \sim \epsilon_c / 2$, i.e.\ not too close to critical, then
$\Delta \phi^2 \sim 1$, so except for $\epsilon_c - \epsilon \ll \epsilon_c$,
the second term dominates the first and we have one real and one imaginary root,
corresponding to the two CSs. Otherwise, we can identify that for
\begin{align*}
    \Delta \phi &\leq \eta^3 \sin I \cos^2 I,\\
    \epsilon_c - \epsilon &\leq \frac{\eta^7 \sin^3 I \cos^4 I}{2},
\end{align*}
that CS2 becomes unstable. We can also numerically validate this via
\lstinline{99_misc/0_stab.py}.

\subsection{Dissipating Disk: $\theta_f$ Distribution}

Consider if $\eta$ crosses $\eta_c$ with $\eta(t = 0) > \eta_c$, and there is no
dissipation, only variation of $\eta$. The area contained within the separatrix
when it first forms is given in Ward \& Hamilton Paper I\@:
\begin{equation}
    A_{crit} = 4\pi \s*{1 - \p*{1 + \tan^{2/3}I}^{-3/2}}.
\end{equation}
We are interested in states that start nearly aligned w/ the disk, so in Zone II
by our designation. Then, once three states appear, Zone II continues to shrink
until separatrix crossing occurs. At the time of separatrix crossing, the
trajectory is ejected into either Zone I/III (if Zone III is shrinking, then
exclusively Zone I\@; can analytically state using Henrard \& Murigande 1987).
The possible values for the new action of the ejected state are either $A_2 +
A_3 - 2\pi$ or $A_3 - 2\pi$, when integrating $\mu\;\mathrm{d}\phi$. From here,
the states will precess uniformly at some $\theta_f = A / 2\pi$.
